---
title:  'Coding Challenge #4'
subtitle: 'SWORK 881'
author: 
- name: Student -  [Talin Gharibian]
output: html_document
---

#install new packages nnet, MASS, quantreg, splines

```{r}
#install.packages("nnet", dependencies = T)
#install.packages("MASS", dependencies = T)
#install.packages("quantreg", dependencies = T)
#install.packages("splines", dependencies = T)
```


## load libraries
```{r}
library(tidyverse)
library(tidymodels)
library(nnet)
library(MASS)
library(quantreg)
library(splines)
```

## load dataset
```{r}
data.01<-read_csv("Data/LHS_exercise_02102022.csv")
```


#data wrangling

We will be creating new variables: 
`anxiety` = clincally significant symptoms of anxiety; score 3 or more on PHQ-4 items 1 & 2
`depression` = clinically significant symptoms of depression; score of 3 or mroe on PHQ-4 items 3 & 4 
`distress` = continuous PHQ-4 score; higher score indicate greater distress

`distress_type` = multinomial variable: 0- none; 1- anxiety; 2- dep; 3- anxiety & dep --> come up with a new variable called distress type that has 4 possible values

(hint: use `str_c(var1, '', var2)` to concatenate two values by rows)
`distress_cat` = categorizing distress score into min:4 ~ 'low'; 5:8 ~ 'mid'; 8:max ~ 'high'
(hint: use `%in%`)
`age_sd` = standardized `age` variable


```{r}
data.02 <- data.01%>%
  mutate(anxiety = case_when(phq1_1 + phq2_1 >= 3 ~ 1,
                           TRUE ~ 0),
         depression = case_when(phq3_1 + phq4_1 >= 3 ~ 1,
                           TRUE ~ 0),
         distress = phq1_1 + phq2_1 + phq3_1 + phq4_1,
         distress_type_pre = str_c(anxiety,depression),
         distress_type = case_when(distress_type_pre == '00' ~ 0,
                                   distress_type_pre == '10' ~ 1,
                                   distress_type_pre == '01' ~ 2,
                                   distress_type_pre == '11' ~ 3),
         distress_cat = case_when(distress <= 4 ~ 'low',
                                  distress %in% c(5,6,7,8) ~ 'mid',
                                  distress >= 8 ~ 'high'),
         age_sd = scale(age)
         )


```

```{r}
ftable(data.02$distress_type)
```

Report frequency of `distress_type`

0   1   2   3

322  67  54 182

_hint_: your **none** category must have $n=322$; if you do not have this, check your code again.

```{r}
ftable(data.02$distress_cat)
```

Report frequency of `distress_cat`

 high low mid
             
  111 344 170

_hint_: your **mid** category must have $n=170$; if you do not have this, check your code again.

---
## multinomial logistic regression

Run multinomial regression where DV is `distress_type`. Use `anxiety` as your reference category. 
Use your typical list of independent variables (standardized age, sex, education). 
 
_hint_: use `nnet` package's `multinom` function

```{r}
#first change character to format type:
data.02$distress_type = as.factor(data.02$distress_type)

#relevel
data.02$distress_type <- relevel(data.02$distress_type, ref= 1)

#now we are ready for running multinomial logistic regression

fit.mlr<- multinom(distress_type ~ age_sd + factor(sex) + factor(education), 
                   data=data.02)

#summary(fit.mlr)

tidy(fit.mlr)

```

Please note your interpretation unit and your outcome unit must be the same. If needed, trasnform your output to facilitate  your interpretation.

Focusing on those who reported both _anxiety_ **AND** _depression_ as your outcome level of interest, provide your interpretation below:
> [I'm not sure.]


---
## ordinal logistic regression

Run ordinal regression where DV is `distress_cat`. Use your typical list of independent variables (standardized age, sex, education).

```{r}
#make sure your variable is a factor type

data.02$distress_cat_factor <- as.factor(data.02$distress_cat)

fit.olr<- polr(distress_cat_factor~
                 age_sd+
                 factor(sex), 
               data=data.02,
               Hess = TRUE)

#Hess option is used to get standard errors

tidy(fit.olr, conf.int = T, 
     p.values = T)


```
To the best of your knowledge, provide explanation of the result. 

> [I'm not sure]


---
## splines!

Run linear regression where DV is `distress`. For your standardized age variable, locate knot points at `-0.5`, `0`, and `0.5`. 

```{r}
fit.01 <- lm(distress ~ age_sd + factor(sex) + factor(education), 
                   data=data.02)

summary(fit.01)
```

Provide your interpretation below:
> [Age and sex are significant in predicting distress. An increase in age is negatively associated with distress. 


Run logistic regression where DV is `depression`. For your standardized age variable, locate knot points at  `-1`, `0`, and `1`. 

```{r}
fit.02 <- glm(depression ~ age_sd + factor(sex) + factor(education), 
                   family=binomial (link="logit"),
              data=data.02)
summary(fit.02)
```
Provide your interpretation below:
> [Standardized age is significant in predicting depression score. An increase in age is negatively associated with depression score.]
